The watershed transform in ITK - discussion and new developments

### 2006-04-23 article/Article.tex
add algorithms 2 and 3 pseudocode
darcs-hash:20060423124020-2fc9d-c3d5616c7aa5a6eabd1c501041bfcee086aaa0a4.gz

\subsubsection{Algorithm 2}
Using ordered queues.
\begin{itemize}
\item initialize the queue with edges of the regional minima. 
The edges are pixels that are part of a regional minima with
neighbours that are background.
\item while (!empty(queue))
   \begin{itemize}
	\item Take the pixel $p$ from the head of the queue
        \item for all pixels $q$ that are steepest upper neighbors of $p$; do
         \item  if $q$ is unlabeled then

		place $q$ in the queue

\end{itemize}

@@

\subsubsection{Algorithm 3}
Also using ordered queues.
\begin{itemize}
\item initialize the queue with background pixels that have 
      neighbors that are members of a regional minima. In practice
      this is done by checking all regional minima pixels to see
      whether they have background neighbors. The background pixels
      are placed on the queue and the label of the regional minima
      associated with them.
\item while (!empty(queue)
   \begin{itemize}
	\item Take the pixel $p$ from the head of the queue

         if $p$ is unlabelled;

           apply the associated label

         else

           check whether $p$ is a watershed pixel. $p$ is a watershed
           pixel if there is more than one label in the neighborhood.

        \item for all pixels $q$ that are steepest upper neighbors of $p$; do

	if $q$ is not in the queue;

           place in queue, associated with label($p$)

           mark as queued
   \end{itemize}
\end{itemize}

>>>>

\subsubsection{Algorithm 2}

Using ordered queues.

This algorithm does not mark the watershed lines.

@@

This algorithm take a grayscale image $I$ and a marker image $M$
which contains a set of markers with different labels. It can be
splitted in to steps:
\begin{enumerate}
  \item copy the marker image $M$ to the output image, and put
all the marker pixels $p$ with background neighbors in the
ordered queue, with the priority $I[p]$.
  \item get a pixel from the ordered queue at the highest
priority level, and propagate the label of $p$ to all the unlabeled
neighbors of $p$. The unlabeled neighbors $q$ are added in the
ordered queue with priority $I[q]$, if $I[q]$ is greater than
the priority level of $p$, or with the priority of $p$ otherwise.
\end{enumerate}

The detailed algorithm is shown in Figure \ref{BeucherAlgorithm}.

\begin{figure}[htbp]
\centering
\small
\begin{pseudocode}[framebox]{BeucherWatershed}{I,M}
\COMMENT{initialization stage} \\
queue \GETS \emptyset \\
\FOREACH p \in M \DO
\BEGIN
  \COMMENT{copy the marker image to the output image}\\
  O[p] \GETS M[p] \\
  \COMMENT{put the edges of the marker in the queue}\\
  \IF M[p] \neq WSLABEL
  \THEN
  \BEGIN
    bgNeighbor \GETS \FALSE \\
    \FOREACH q \in \CALL{N}{p} \DO 
    \BEGIN
      \IF M[q] = WSLABEL
        \THEN bgNeighbor \GETS \TRUE \\
    \END\\
    \IF bgNeighbor \STMTNUM{5em}{algo:Beucher:bgNeighbor}
      \THEN
      queue[I[p]] \GETS queue[I[p]] \cup \{p\} \\
  \END
\END
\\
\\
\COMMENT{flooding stage} \\
\WHILE queue \neq \emptyset \DO
\BEGIN
  v \GETS \CALL{first}{queue}\\
  \WHILE queue[v] \neq \emptyset \DO
  \BEGIN
    p \GETS \CALL{first}{queue[v]} \\
    \FOREACH q \in \CALL{N}{p} \DO
    \BEGIN
    \IF O[q] = WSLABEL
    \THEN
    \BEGIN
      O[q] \GETS O[p] \\
      v \GETS \CALL{MAX}{I[q], v} \STMTNUM{5em}{algo:Beucher:max} \\
      queue[v] \GETS queue[v] \cup \{q\} \\
    \END
    \END
  \END
\END
\end{pseudocode}
\caption{Simple watershed algorithm.\label{BeucherAlgorithm} $FIRST(queue)$ extract the highest priority element from the queue. $N(p)$ return the list of p's neighbors. $MAX(a,b)$ returns the maximum of $a$, and $b$. $WSLABEL$ is the label used to define background and watershed pixels (usally $0$). The lower completion is done by the statement (\ref{algo:Beucher:max}). Determine if a marker pixel before adding it in the queue (\ref{algo:Beucher:bgNeighbor}) is not required, but helps to decrease memory usage in case of large markers.}
\end{figure}

% \begin{itemize}
% \item initialize the queue with edges of the regional minima. 
% The edges are pixels that are part of a regional minima with
% neighbours that are background.
% 
% \item while (!empty(queue))
%    \begin{itemize}
% 	\item Take the pixel $p$ from the head of the queue
%         \item for all pixels $q$ that are steepest upper neighbors of $p$; do
%          \item  if $q$ is unlabeled then 
% 
%      		propagate label$(p)$ to label$(q)$;
% 	
% 		place $q$ in the queue
% 
%                 else if (label$(q) \ne \mbox{WS})$ and  (label$(q) \ne \mbox{label}(p)$ then
% 
%                 label($q$) = WS;
% 	\end{itemize}
% 
% \end{itemize}

\subsubsection{Algorithm 3}

Also using ordered queues.

This algorithm take a grayscale image $I$ and a marker image $M$
which contains a set of markers with different labels. It can be
splitted in to steps:
\begin{enumerate}
  \item copy the marker image $M$ to the output image, and put
all the background pixels $p$ with markers in there neighbors in the
ordered queue, with the priority $I[p]$, and mark $p$ as already in
the queue. All the marker pixels are marked as already in the queue.
  \item get a pixel $p$ from the ordered queue at the highest
priority level. If the neighborhood contains only the same label
or watershed pixels, mark $p$ with that label. The neighbors $q$ 
of $p$ which are not already in the queue are added to the queue 
with priority $I[q]$, if $I[q]$ is greater than the priority level
of $p$, or with the priority of $p$ otherwise.
\end{enumerate}

The detailed algorithm is shown in Figure \ref{MeyerAlgorithm}.

Contrary to the previous algorithm, the label in the output image
can't be used to determine if the pixel has already been added to
the queue, and so we need to maintain this status by another way.
It is the role of $S$ in Figure \ref{MeyerAlgorithm}. We'll discuss
in ??? how this status can be implemented.

This algorithm mark the watershed lines.

It must be noticed that this algorithm never propagate watershed
lines, and so produce very thin watershed lines.

\begin{figure}[htbp]
\centering
\small
\begin{pseudocode}[framebox]{MeyerWatershed}{I,M}
\COMMENT{initialization stage} \\
\FOREACH p \in M \DO
  S[p] \GETS \FALSE \\
queue \GETS \emptyset \\
\FOREACH p \in M \DO
\BEGIN
  \COMMENT{copy the marker image to the output image}\\
  O[p] \GETS M[p] \\
  \IF M[p] \neq WSLABEL
  \THEN
  \BEGIN
    S[p] \GETS \TRUE \\
    \COMMENT{add all the neighbors of $p$ with WSLABEL to the queue}\\
    \FOREACH q \in \CALL{N}{p} \DO
    \BEGIN
      \IF \neg S[q] \AND M[q] = WSLABEL 
      \THEN
      \BEGIN
        S[q] \GETS \TRUE \\
        queue[I[q]] \GETS queue[I[q]] \cup \{q\} \\
      \END
    \END
  \END
\END
\\
\\
\COMMENT{flooding stage} \\
\WHILE queue \neq \emptyset \DO
\BEGIN
  v \GETS \CALL{first}{queue}\\
  \WHILE queue[v] \neq \emptyset \DO
  \BEGIN
    p \GETS \CALL{first}{queue[v]} \\
    \COMMENT{found the label of $p$ and determine if it is a watershed pixel} \\
    label \GETS WSLABEL \\
    watershed \GETS \FALSE \\
    \FOREACH q \in \CALL{N}{p} \DO
    \BEGIN
      \IF O[q] \neq WSLABEL \AND \neg watershed
      \THEN
        \BEGIN
        \IF label \neq WSLABEL \AND O[q] \neq label \\
        \THEN
          watershed \GETS \TRUE \\
        \ELSE
          label \GETS O[q] \\
        \END
    \END \\
    \COMMENT{if $p$ is a watershed pixel, don't propagate this pixel} \\
    \IF \neg watershed
    \THEN
      \BEGIN
        \COMMENT{assign the label value to the current pixel} \\
        O[p] \GETS label \\
        \COMMENT{put all the neighbors of p in the queue} \\
        \FOREACH q \in \CALL{N}{p} \DO
        \BEGIN
          \IF \neg S[q]
          \THEN
            \BEGIN
            S[q] \GETS \TRUE \\
            h \GETS \CALL{MAX}{I[q], v} \STMTNUM{5em}{algo:Meyer:max} \\
            queue[h] \GETS queue[h] \cup \{q\} \\
            \END
        \END
      \END
  \END
\END
\end{pseudocode}
\caption{Watershed algorithm with watershed lines.\label{MeyerAlgorithm} $FIRST(queue)$ extract the highest priority element from the queue. $N(p)$ return the list of p's neighbors. $MAX(a,b)$ returns the maximum of $a$, and $b$. $WSLABEL$ is the label used to define background and watershed pixels (usally $0$). The lower completion is done by the statement (\ref{algo:Meyer:max})}
\end{figure}

% \begin{itemize}
% \item initialize the queue with background pixels that have 
%       neighbors that are members of a regional minima. In practice
%       this is done by checking all regional minima pixels to see
%       whether they have background neighbors. The background pixels
%       are placed on the queue and the label of the regional minima
%       associated with them.
% \item while (!empty(queue)
%    \begin{itemize}
% 	\item Take the pixel $p$ from the head of the queue
% 	
%          if $p$ is unlabelled;
% 
%            apply the associated label
% 
%          else
%  
%            check whether $p$ is a watershed pixel. $p$ is a watershed
%            pixel if there is more than one label in the neighborhood.
% 
%         \item for all pixels $q$ that are steepest upper neighbors of $p$; do
% 	
% 	if $q$ is not in the queue;
% 
%            place in queue, associated with label($p$)
%         
%            mark as queued
%    \end{itemize}
% \end{itemize}
